---
title: "Untitled"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
set.seed(123)
N <- 5000

T_time <- 50
#take the forced expiratory volume within response (FEV1)

beta <- 2 #as interception
beta_smoking <- -0.5 #smoking coefficient
beta_treatment <- 0.3 #
beta_age <- -0.03
beta_BMI <- -0.05
sigma_b <- 1.2 #random effect
sigma_e <- 1.1 #measurement error

age <- rnorm(N, mean = 50, sd = 10)
BMI <- rnorm(N, mean = 25, sd = 3)

treatment <- rbinom(N, 1, 0.5)

```

```{r}


library(dplyr)
library(geepack)

set.seed(123) 

# ------------------------------
# 1. 参数设定
# ------------------------------
N_all <- 1000        # Sample size in total
T_time <- 5          # Five years observation
sigma_b <- 0.8       # random effects in participants
sigma_e <- 1.0       # error in measurement 

# beta
beta0 <- 2.4                 # Intercept
beta_smoking <- -0.15        #  About 0.15 L in FEV when smoking
beta_treat <- 0.10           # Treatment applied
beta_age <- -0.02            # 
beta_gender <- 0.30          # Male is higher 
beta_BMI <- -0.03            # 
beta_weight <- 0.01          # 
beta_packyears <- -0.02      # Year of smoking
beta_cigsday <- -0.005       # 
beta_sbp <- -0.004           # SBP
#beta_baseFEV1 <- 0.8         # baseline FEV1 效应

# ------------------------------
# 2. Baseline factors of each participants
# ------------------------------
id <- 1:N_all
age <- rnorm(N_all, mean = 50, sd = 10)
gender <- rbinom(N_all, 1, 0.5)              # 0=女, 1=男
BMI <- rnorm(N_all, mean = 25, sd = 3)
weight <- rnorm(N_all, mean = 70, sd = 10)
pack_years <- rpois(N_all, lambda = 20)      # 吸烟年数，Poisson(20)做示例
cigs_per_day <- rpois(N_all, lambda = 15)    # 每日吸烟量，Poisson(15)示例
sbp <- rnorm(N_all, mean = 120, sd = 15)     # 收缩压
#baseFEV1 <- rnorm(N_all, mean = 2.3, sd = 0.4) # 基线 FEV1
treatment <- rbinom(N_all, 1, 0.5)           

#Random effects of participants
b_i <- rnorm(N_all, mean = 0, sd = sigma_b)

#Baseline factors
df_subjects <- data.frame(
  id, age, gender, BMI, weight,
  pack_years, cigs_per_day, sbp,
  treatment, b_i
)


dat_all <- data.frame()  

for(i in seq_len(N_all)){
  for(t in 1:T_time){
    smoking_it <- rbinom(1, 1, 0.4)# The smoking status, assume 40% of the sample smoke
    
    mu <- beta0 +
          beta_smoking     * smoking_it +
          beta_treat       * df_subjects$treatment[i] +
          beta_age         * df_subjects$age[i] +
          beta_gender      * df_subjects$gender[i] +
          beta_BMI         * df_subjects$BMI[i] +
          beta_weight      * df_subjects$weight[i] +
          beta_packyears   * df_subjects$pack_years[i] +
          beta_cigsday     * df_subjects$cigs_per_day[i] +
          beta_sbp         * df_subjects$sbp[i] +
          #beta_baseFEV1    * df_subjects$baseFEV1[i] +
          df_subjects$b_i[i]   # Random effect in each participant
    
    # Random effect in of each FEV1
    FEV1_it <- rnorm(1, mean = mu, sd = sigma_e)
    
    dat_all <- rbind(dat_all, data.frame(
      id = i,
      time = t,
      smoking = smoking_it,
      treatment = df_subjects$treatment[i],
      age = df_subjects$age[i],
      gender = df_subjects$gender[i],
      BMI = df_subjects$BMI[i],
      weight = df_subjects$weight[i],
      pack_years = df_subjects$pack_years[i],
      cigs_per_day = df_subjects$cigs_per_day[i],
      sbp = df_subjects$sbp[i],
      #baseFEV1 = df_subjects$baseFEV1[i],
      FEV1 = FEV1_it
    ))
  }
}

dat_all$id <- factor(dat_all$id)

#split into 100, 500, 1000 as in the paper
dat_100  <- dat_all[dat_all$id %in% 1:100, ]
dat_500  <- dat_all[dat_all$id %in% 1:500, ]
dat_1000 <- dat_all[dat_all$id %in% 1:1000, ]

```

```{r}
dat_all
#count(dat_all)
unique(dat_all)
```
```{r}
dat_all[dat_all$time < 4, ]
```

```{r}
library(dplyr)

df_year <- dat_all %>% filter(time == 2)
df_year
```

A linear regression model was fitted at each of the five visit years with the following covariates: the current-year smoking-status, treatment group and some baseline characteristics (age, gender, body mass index, body weight, smok ing pack-years, cigarettes smoked per day, systolic blood pressure and FEV1). 

it is confirmed that quitting smoking is associated with an increase of FEVi by about 0.09 to 0.19 
μ = β0 + β1 * smoking + β2 ×treatment + β3×age+β4×BMI+⋯+subject_effect

Average values in healthy patients aged 20-60 range from 4.5 to 3.5 liters in males and from 3.25 to 2.5 liters in females
Sustained smoking cessation substantially reduces the smoking-related accelerated rate of decline in forced expiratory volume in one second


```{r}

cross_sec_res <- function(data_cross_sec){
  output <- data.frame(
    Year = integer(), Smoking_est = numeric(), Smoking_se = numeric()
  )
  for(t in 1:5){
    df_year <- data_cross_sec %>% filter(time == t)#t=1,2,3,4,5
    fit <- lm(FEV1 ~ smoking + treatment + age + gender + BMI +weight +
                pack_years + cigs_per_day + sbp, data = df_year)
    
    est <- coef(summary(fit))["smoking","Estimate"]
    se  <- coef(summary(fit))["smoking","Std. Error"] #Extract the lm() parameters & coefficient 
    
    output <- rbind(output, data.frame(Year=t, Smoking_est=est, Smoking_se=se))
    print(coef(summary(fit)))
  }
  output
}

cs_res_100 <- cross_sec_res(dat_100)
cs_res_100

```
```{r}
cs_res_500 <- cross_sec_res(dat_500)
cs_res_500
cs_res_1000 <- cross_sec_res(dat_1000)
cs_res_1000

```


```{r}
cbind(cs_res_100, cs_res_500,cs_res_1000)
```


```{r}
gee_res <- function(data_gee){
  
  fit_indep <- geeglm(FEV1 ~ smoking + treatment + age + gender + BMI + weight + pack_years + cigs_per_day + sbp, id = id, data = data_gee, family = gaussian, corstr = "independence")
  fit_cs <- geeglm(FEV1 ~ smoking + treatment + age + gender + BMI + weight + pack_years + cigs_per_day + sbp, id = id, data = data_gee, family = gaussian, corstr = "exchangeable")
  fit_ar1 <- geeglm(FEV1 ~ smoking + treatment + age + gender + BMI + weight + pack_years + cigs_per_day + sbp, id = id, data = data_gee, family = gaussian, corstr = "ar1")
  #print(coef(summary(fit)))
  extract_coef <- function(fit, value){
    print(summary(fit)$coefficients)
    coef <- summary(fit)$coefficients
    c(Estimate = coef[value, "Estimate"], SE = coef[value, "Std.err"])
  }
  
  smoke_ind <- extract_coef(fit_indep, "smoking")
  smoke_cs <- extract_coef(fit_cs, "smoking")
  smoke_ar1 <- extract_coef(fit_ar1, "smoking")
  
  output<- data.frame(
    Model = c("indep", "CS", "AR_1"),
    Smoking_est = c(smoke_ind["Estimate"], smoke_cs["Estimate"], smoke_ar1["Estimate"] ),
    Smoking_se = c(smoke_ind["SE"], smoke_cs["SE"], smoke_ar1["SE"] )
  )
  output
}

gee_res_100 <- gee_res(dat_100)
gee_res_100
```
```{r}

gee_res_500 <- gee_res(dat_500)
gee_res_500

```

```{r}
gee_res_all <- gee_res(dat_all)
gee_res_all
```

```{r}

#Implement PMSE
pmse_boot <- function(data, indices, corstructure_type) {
  # indices: bootstrap 
  subjects <- unique(data$id)
  boot_subjects <- subjects[indices]
  # retain one participant
  boot_data <- data[data$id %in% boot_subjects, ]
  
  # 
  fit <- geeglm(FEV1 ~ smoking + treatment + age + gender + BMI + weight +
                  pack_years + cigs_per_day + sbp + baseFEV1,
                id = id, data = boot_data, family = gaussian, corstr = corstructure_type)
  
  pred <- predict(fit, newdata = data)
  # The SE of the error square?
  se <- (data$FEV1 - pred)^2
  
  return(mean(se))
}

B <- 50
n_subj <- length(unique(dat_100$id)) 
get_pmse <- function(corstructure_type, data) {
  boot_pmse <- replicate(B, {
    boot_indices <- sample(1:n_subj, n_subj, replace = TRUE)
    pmse_boot(data, boot_indices, corstr_type)
  })
  mean(boot_pmse)
}

pmse_indp <- get_pmse("independence", dat_100)




print(pmse_results)



```













